Modelling the Trojan Asteroids

### 2020-04-27 Report/draft1.tex
Final submission

%\author{
	%\authorstyle{Christopher Gallagher}}
\author{\authorstyle{Christopher Gallagher} %Candidate 6952R}
	\institution{University of Cambridge}}
% Example of a one line author/institution relationship
%\author{\newauthor{John Marston} \newinstitution{Universidad Nacional Autónoma de México, Mexico City, Mexico}}

@@

\lettrineabstract{Trojan asteroids orbit the Sun in resonance with Jupiter at Lagrange points,
	and have their own stable orbits about these points. This can be treated as a restricted three-body problem, which is numerically evaluated in this report using a stiff Radau method to solve the system of coupled differential equations. Without perturbations, a maximal deviation of 4.68\pmb{$\times$}10\pmb{$^{-13}$} AU  from the Lagrange points was observed over 1000 orbits (11,852 years). The application of perturbations to these asteroids demonstrated the predicted existence of stable tadpole and horseshoe orbits, and suggested that wander from the Lagrange point depends quadratically on perturbation size. Such orbits were observed to be stable up to a planetary mass of 0.04 \pmb{$M_{\odot}$}, also in agreement with theoretical predictions. Word Count: XXXX}

@@

\section{Methodology}
This system of coupled first-order ordinary differential equations (ODEs) was solved using the scipy \textit{solve\_ivp} function. The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years. Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in Earth years and mass in multiples of the solar mass, to prevent floating point overflows due to the magnitude of the quantities considered in SI units.

@@

\subsection{Wander Analysis} \label{perturbed}
Wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin. The perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively) were considered separately. By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).

\begin{figure}[ht]
	\centering
	\begin{subfigure}{.45\textwidth}
		\centering
		% include first image
		\includegraphics[width=\linewidth]{Figures/wanderagainstradialpertubation2}  
		\caption{Radial}
		\label{fig:wander_rad}
	\end{subfigure}
	\hfill %% useful if width of each figure is less the .5\textwidth
	\begin{subfigure}{.45\textwidth}
		\centering
		% include second image
		\includegraphics[width=\linewidth]{Figures/wanderagainsttangentialpertubation2}  
		\caption{Tangential}
		\label{fig:wander_tan}
	\end{subfigure}
	\caption{Wander from an initial perturbation with maximum relative magnitude of 0.1\% of the displacement from the origin, given in radial and tangential directions.}
	\label{fig:wanderplots}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/z_perturbations5c}
	\caption{Variation of wander with increasing Z perturbations. A strong linear component is observed initially in the un-perturbed case, but then tends towards the perturbed quadratic case, while the linear offsets result from the initial perturbation in the XY plane.}
	\label{fig:zwander}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/z_perturb_orbit}
	\caption{Orbit of Greek asteroids following a 0.2 AU perturbation in the positive Z direction, over 12 orbital periods. Note the initial outwards radial displacement despite the lack of perturbation in the XY plane, and subsequent tadpole-like orbit.}
	\label{fig:zorbit}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_q7}
	\caption{Variation of maximum wander with planetary mass, under a 0.001 AU radial perturbation and 0.001 AU/year tangential boost. Note the tendency }
	\label{fig:planetmass}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p6}
	\caption{Variation of maximum wander with planetary mass, under a 0.01 AU radial perturbation and 0.01 AU/year tangential boost. Note the orbit stability up to $ M_{p} = 0.04 M_{s}$, and instability beyond this.}
	\label{fig:planetmass2}
\end{figure}

>>>>

%\author{
	%\authorstyle{Christopher Gallagher}}
\author{\authorstyle{Candidate 6952R}
	\institution{University of Cambridge}}
% Example of a one line author/institution relationship
%\author{\newauthor{John Marston} \newinstitution{Universidad Nacional Autónoma de México, Mexico City, Mexico}}

@@

\lettrineabstract{Trojan asteroids orbit the Sun in resonance with Jupiter at Lagrange points,
	and have their own stable orbits about these points. This can be treated as a restricted three-body problem, which is numerically evaluated in this report using a stiff Radau method to solve the system of coupled differential equations. Without perturbations, a maximal deviation of 4.68\pmb{$\times$}10\pmb{$^{-13}$} AU  from the Lagrange points was observed over 1000 orbits (11,852 years). The application of perturbations to these asteroids demonstrated the predicted existence of stable tadpole and horseshoe orbits, and suggested that wander from the Lagrange point depends quadratically on perturbation size. Such orbits were observed to be stable up to a planetary mass of 0.04 \pmb{$M_{\odot}$}, also in agreement with theoretical predictions. Word Count: 2782}

@@

\section{Methodology}
This system of coupled first-order ordinary differential equations (ODEs) was solved using the \textit{solve\_ivp} function from \textit{scipy}. The time span was taken as 100 orbits (with 100 points sampled per orbit) unless otherwise stated; this corresponds to 1185 Earth years. Rescaled solar system units are used for mathematical ease, so distances are measured in astronomical units (AU), time in Earth years and mass in multiples of the solar mass, to prevent floating point overflows due to the magnitude of the quantities considered in SI units.

@@

\subsection{Wander Analysis} \label{perturbed}

\begin{figure}[hb!]
	\centering
	\begin{subfigure}{.45\textwidth}
		\centering
		% include first image
		\includegraphics[width=\linewidth]{Figures/wanderagainstradialpertubation2}  
		\caption{Radial}
		\label{fig:wander_rad}
	\end{subfigure}
	\hfill %% useful if width of each figure is less the .5\textwidth
	\begin{subfigure}{.45\textwidth}
		\centering
		% include second image
		\includegraphics[width=\linewidth]{Figures/wanderagainsttangentialpertubation2}  
		\caption{Tangential}
		\label{fig:wander_tan}
	\end{subfigure}
	\caption{Wander from an initial perturbation with maximum relative magnitude of 0.1\% of the displacement from the origin, given in radial and tangential directions.}
	\label{fig:wanderplots}
\end{figure}

Wander from the initial point was calculated for random perturbations with a maximum magnitude of 1\% of the displacement from the origin. The perturbation components parallel and perpendicular to the position vector from the CoM (hereafter referred to as radial and tangential components respectively) were considered separately. By considering these perturbations across position space, it was clear that the wander is fully determined by the radial component, with no tangential dependence (as shown in Figure \ref{fig:wanderplots}).

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/z_perturbations5c}
	\caption{Variation of wander with increasing Z perturbations. A strong linear component is observed initially in the un-perturbed case, but then tends towards the perturbed quadratic case, while the linear offsets result from the initial perturbation in the XY plane.}
	\label{fig:zwander}
\end{figure}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/z_perturb_orbit}
	\caption{Orbit of Greek asteroids following a 0.2 AU perturbation in the positive Z direction, over 12 orbital periods. Note the initial outwards radial displacement despite the lack of perturbation in the XY plane, and subsequent tadpole-like orbit.}
	\label{fig:zorbit}
\end{figure}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_q7}
	\caption{Variation of maximum wander with planetary mass, under a 0.001 AU radial perturbation and 0.001 AU/year tangential boost. Note the tendency }
	\label{fig:planetmass}
\end{figure}

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{Figures/wanderwithplanetmass_p6}
	\caption{Variation of maximum wander with planetary mass, under a 0.01 AU radial perturbation and 0.01 AU/year tangential boost. Note the orbit stability up to $ M_{p} = 0.04 M_{s}$, and instability beyond this.}
	\label{fig:planetmass2}
\end{figure}
